### 2024-09-12 article/Article.tex
updated article, pdfs, README

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{TWINKLE: An open-source two-photon microscope for teaching and research}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Manuel Schottdorf\textsuperscript{1,*},
P. Dylan Rich\textsuperscript{1},
E. Mika Diamanti\textsuperscript{1}, 
Albert Lin\textsuperscript{1,4}, 
Sina Tafazoli\textsuperscript{1},
Edward H. Nieh\textsuperscript{1,2},
Stephan Y. Thiberge\textsuperscript{1,3*}
\\
\bigskip
\textbf{1} Princeton Neuroscience Institute, Princeton University, Princeton, NJ, USA\\
\textbf{2} School of Medicine, University of Virginia, Charlottesville, VA, USA \\
\textbf{3} Bezos Center for Neural Circuit Dynamics, Princeton University, Princeton, NJ, USA\\
\textbf{4} Center for the Physics of Biological Function, Princeton University, Princeton, NJ, USA\\
\bigskip

@@

\section*{Materials and methods}
In two-photon laser-scanning microscopy\cite{Denk1990}, laser light is focused to a small volume. The product of the interaction of light and sample, fluorescent light, is then measured with a sensitive light detector. Moving the volume of light through the sample and combining the detected fluorescence from different $x$ and $y$ positions produces a picture. The key to two-photon microscopy is the light source: A femtosecond-pulsed infrared laser restricts the volume in which fluorescence occurs to a tiny region in the sample with sufficiently high photon density for the near-simultaneous absorption of two photons by a fluorescent molecule \cite{Denk1990, Svoboda1997, So2000, Helmchen2005}.\newline
Here, we design an open and high-performing microscope that uses these principles. As such we have made all optical and mechanical designs, electronics, bill-of-materials, CAD assembly, testing results, and relevant schematics open to the public. Excluding laser and optical table, the cost is around US-\$ 110,000 (2024). A narrative of the design and high-level view is presented here. For further technical details, and complete built documentation, we refer the reader to \nameref{S1_Appendix}, and the repository at \url{https://github.com/BrainCOGS/Microscope}.

\subsection*{Design specifications}
To summarize our design goals, we aimed for a mechanical and optical assembly using as many off-the-shelf components as possible, with only few custom aluminium parts that can be machined in any university machine shop. Our design lacks a microscope body, which is  advantageous for two reasons: it aids the use as a teaching instrument, and facilitates adaptation for \invivo imaging because it provides space for the organism and ancillary hardware. We also avoid any custom optics to aid transparency of the function of the optical system. Our system is also made cost-effective, in part, by the availability of femtosecond laser systems based on fiber technology. Operating at a fixed wavelength, these systems come at the fraction of the cost of a more traditional tunable Ti:Sapphire laser \cite{Bueno2019}.\newline
We aimed for (1) $\gtrapprox20\text{ cm}$ of free space around the objective in all directions to aid integration of the microscope into custom behavior boxes, or sample container, and (2) resolution limited imaging of at least $\approx 700\times 700\text{ \textmu m}^2$ field of view which is typical for imaging brain tissue at cellular resolution. (3) Compared to previous open designs\cite{Rosenegger2014, Mayrhofer2015}, our system can make use of the large back aperture of modern long-working-distance objectives to collect more fluorescence and (4) uses GaAsP photomultiplier tubes to maximize signal to noise.\newline
An overview of the microscope is shown in Fig.~\ref{fig1}. Fig.~\ref{fig1}A shows the system as a cartoon. Full details will be provided below. In short, we chose a Spark Alcor 920 nm pulsed femtosecond laser, whose beam, after $2\times$ magnification with a telescope, reaches the scanning mirrors. The scanning mirrors are geometrically close together in the focal plane of the scan lens. This simplifies the design, but leads to minor distortions across the field of view (a few percent at low magnifications, see below). The combined scan and tube lenses magnify the beam further to fill the back aperture of the objective. Fluorescence from the specimen is then collected and focused on the aperture of two GaAsP photomultiplier tubes with suitable color filters. Fig.~\ref{fig1}B shows the CAD design of the entire assembly with important parts of the optical path labelled. Fig.~\ref{fig1}C shows an annotated picture of the microscope head. Notice that compared to the cartoon in Fig.~\ref{fig1}A, the optical path is folded with several mirrors. These mirrors significantly simplify alignment, and make the design more compact.

@@

\subsection*{Assembly on the table}
Various components are used between the laser and the microscope head to condition the beam, and  for intensity control, Fig.~\ref{fig1}B. Out of the laser head, the beam first encounters a waveplate (AHWP10M-980, Thorlabs). This is a birefringent crystal that can rotate the polarization of the laser. Next, the beam travels through a polarizing beamsplitter cube (PBS103, Thorlabs). This splits the beam into two beams whose relative intensity can be adjusted by the orientation of the waveplate. We set the orientation of the waveplate to split the beam into two beams with 50\% power each to feed a second experimental setup. In the CAD drawing, this beam is sent to a beam block (LB1, Thorlabs). Next, the beam passes through an electro-optic modulator, a Pockels cell (350-80-LA-02, Conoptics). This device allows rapid and electronic control of the laser's power level. It uses a crystal whose refractive index is controlled by an external electric field, combined with a polarizer. This can be thought of as a voltage-controlled wave plate, in which the electric field controls how much light travels through the polarizer. Next, the beam travels through an open mechanical shutter (LS6S2ZM1, Vincent Associates). When closed, the reflection from the shutter is used for calibrating the Pockels cell with a photodiode (PDA36A2, Thorlabs). Following the shutter, the beam travels though a $2\times$ telescope (GBE02-B, Thorlabs). This magnifies the beam diameter by a factor of 2. Past the telescope, the beam is then reflected off several mirrors in a periscope configuration before entering a custom aluminium box housing, mounted to a 95 mm optical rail, that houses the two scanning mirrors (CRS8K/6215H scanning mirror Set, Novanta), shown in blue in  Fig.~\ref{fig1}C. Based on their angles, the two mirrors deflect the beam in x and y position. The beam then travels through scan lens (green) and tube lens (red), the collection optics (yellow) and finally the objective (cyan). The design of these components is covered in the next paragraphs. The steering mirrors in our system are all silver mirrors from Thorlabs. Silver mirrors contribute minimally to dispersion which is inevitable when light travels through various optical components. Limiting the dispersion makes it possible to correct for it with the laser's built-in dispersion compensation.

@@

We assembled the scan and tube these as lens groups from groups of off-the-shelf parts, using the same elements as in \cite{Yao2023}, and further optimizing lenses and spacing in Zemax (an optics design software). Details are provided below. After optimization, the focal lengths of scan and tube lens were 99 mm and 378 mm respectively, deviating around 1\% from their target values (Zemax simulation at 920 nm).

@@

    \caption{{\bf Optical design of the excitation pathway.} \textbf{A)} Zemax design with the scanning mirror, tube and scan lens groups. Colors indicate light rays produced by different mirror angles of the scanning mirror. Notice how the rays converge in the focal plane. \textbf{B)} CAD design built around the optical design in A, see also Fig.~\ref{fig1}B. \textbf{C)} Cuts through the scan and tube lens assemblies, which are fabricated from off-the-shelf lenses housed in SM2 lens tubes. The specific lenses are stated in the text. \textbf{D)} Simulation of the optical system's two-photon point spread function (PSF-2p) for 0 deg deflection angle, and two example radial sections. \textbf{E)} PSF-2p in xz direction across the full range of deflection angles, and quantification of Petzval field curvature (blue), and radial (purple) and axial (yellow) full-width-at-half-maximum of the PSF-2p. Error bars are resolution limits of the Huygens PSF estimates in sequential mode.}

@@

Fig.~\ref{fig2}A shows the propagation of light rays for various deflection angles of the scanning mirrors converging on points along a line in the focal plane. The Zemax design was then exported into CAD, and the optomechanics assembled around it, see Fig.~\ref{fig2}B. The CAD design and lens groups are highlighted in Fig.~\ref{fig2}B,C (see Fig.~\ref{fig1}C of a photograph of the same part of the instrument). The key property that we optimized when designing the excitation system was an essentially flat, and diffraction limited focal volume across a wide range of deflection angles of the scanning mirrors. Fig.~\ref{fig2}D shows an estimate of the two-photon point spread function (PSF-2p), computed as the square of the Huygens PSF in Zemax. The maximum scan angle of $\pm 10\text{ deg}$ is limited by vignetting of the two inch diameter optics. Analyzing the PSF-2p across optical deflection angles, Fig.~\ref{fig2}E, suggests very small field curvature. In this simulation, the axial resolution was $\delta z = 5.0\text{ \textmu m}$, and the radial resolution $\delta r = 0.75\text{ \textmu m}$ across a wide range of deflection angles.\newline

@@

We incorporated four design principles into our assembly: (1) Fluorescence light must fall approximately perpendicularly on the optical filters. Narrow filters are particularly susceptible to these effect. Although light emitted from the focal plane of an infinity-corrected objective is collimated as it returns from the sample, actual fluorescence weakly diverges due to scattering and diffraction in the material between focal plane and objective (e.g. brain tissue \cite{Zipfel2003, Taddeucci1996}). The detector assembly should therefore be placed as close as possible to the objective where fluorescent light will be nearly collimated. This placement reduces sensitivity to spurious light and allows IR light to travel through the collection optics unimpeded, while fluorescent light is reflected into the optical assembly. This arrangement also has the advantage that possible astigmatism, introduced by curvature of the dichroic in the reflected light, only affects the collection optics which does not have to form an image. (2) The collection lens should be large to maximize the collection of as much fluorescent light as possible. This implies that the collection lens should follow the dichroic mirror as closely as possible. (3) Having the collected photons collimated on the aperture of the photomultiplier tubes reduces the sensitivity of the system on surface inhomogeneities of the detector\cite{Tsai2002, Young2015}. In scanning microscopy, an image is produced as a time series of intensities. It is therefore more important to collect a consistently large portion of fluorescent light, rather than to minimize aberration in the collection optics. We therefore optimized relative illumination as function of the field angle in Zemax. (4) All dichroic mirrors and optical band-pass filters have finite performance. It it therefore advantageous to combine multiple band-pass filters to isolate relevant fluorescence ranges. We aimed to record fluorescence from two well-separated spectral ranges, a green ($525\pm25\text{ nm}$) and a red channel ($600\pm25\text{ nm}$). Beyond an IR block in the collection optics, we chose to only use simple long-pass filters, which means that the red channel is operated in the transmission, and the green channel in the reflection mode (cf. Fig.~\ref{fig1}A).\newline

@@

\subsection*{Mechanical design}
As mounting system of our microscope, we chose a standard 8 inch thick optical table. The design was done in CAD, and the results shown in Fig.~\ref{fig1}B. The aim was a mechanical assembly that uses as few custom parts as possible. We discuss this further in the discussion section. The components in the CAD files can be assembled and aligned by an experienced researcher in a few days. When used for teaching, a careful assembly and alignment is viable with about a week of work. Regarding the footprint, the system can comfortably be built on an $4\times4$ ft area of table. The microscope documented here was setup on a $4\times8$ ft table, shared with a second microscope. The beam-splitting hardware is shown in Fig.~\ref{fig1}B as well.\newline
The custom mechanical components (e.g. the light-tight aluminium housing of the collection optics) were built in the university's machine shop from aluminium stock or Thorlabs parts. Some parts have tight tolerances which suggests manufacturing in a professional machine shop is preferred. If useful for training, this could however be done by students. An example are the aluminium parts that hold the large dichroic mirrors and optical filters in the collection optics. After drilling the apertures for light to pass through, not much metal is left on the part which can lead to warping and distortions.\newline
Finally, we want to emphasize the difference between a theoretically optimal design in Zemax and finite tolerances in real-world optomechanical parts. It is key for a good microscope design to be robust against such errors. In our design, the objective is fixed, while the mirrors, the tube and scan lens assemblies, and the scanning mirrors can move along the optical axis for alignment. This allows the mechanical components sufficient degrees of freedom to optically align all key parts of the microscope. In the supplement, we go through this alignment procedure which allows a user to find the optimal position of the scanning mirrors in the conjugated plane of the back-aperture of the objective.

@@

\subsection*{System performance}
%
Here, we document the performance of the microscope and a few example applications using a Nikon NA 0.8 16$\times$ LWD objective (N16XLWD-PF, Thorlabs). We first determined the size and properties of the field of view with a 100~\textmu m grid (R1L3S3P, Thorlabs), imaged using a Fluorescein in water film through a standard \#1.5 ($\approx170\text{ \textmu m}$ thick) coverslip, see Fig.~\ref{fig5}A. The field of view was $d\approx1.3\text{ mm}$ along the diagonal, $A\approx 1\text{ mm}^2$ in area and the field curvature below the thickness of a thin Fluorescein film $\lesssim 10\text{ \textmu m}$. This is consistent with our earlier Zemax simulations. When measuring the magnification $M$ in \textmu m/px across the field of view, we observed small deviations of $\Delta M \approx 0.1\text{ \textmu m/px}$ in the center when compared to the edges, while the overall range of magnifications across the field of view was tight, see Fig.~\ref{fig5}B. The average magnification is $M=2.77\pm0.04\text{ \textmu m/px}$, which suggests a deviation from a perfect f-Theta system of $\Delta M/M = 0.1/2.8 \approx 4\%$. Next, we imaged a uniform bath of Fluorescein, see Fig.~\ref{fig5}C. This is a combined measure of the quality of excitation and the efficiency of the collection optics across the field of view. Within the central $700\text{ \textmu m} \times 700\text{ \textmu m}$ region, the signal deviated from uniformity within $\approx 10\%-15\%$ ($\pm 1 \text{ standard deviation}$) suggesting a relatively flat field, see Fig.~\ref{fig5}D, E. Following these measurements, we imaged a bead sample of 0.2\microns diameter in 1\% Agarose (Dragon Green beads; Bangs Laboratories) to measure axial and radial PSFs. The sample is shown in Fig.~\ref{fig5}F. Averaging across the $N=38$ beads in this volume, we measured a radially symmetric full-width-at-half-maximum of $\delta r\approx760\pm30\text{ nm}$, see Fig.~\ref{fig5}G,H, and axially $\delta z \approx 5.4\pm0.9\text{ \textmu m}$, see Fig.~\ref{fig5}I. The theoretical resolution limits\cite{Tsai2002} for our system, underfilled to $\text{NA}=0.7$ at $\lambda=920\text{ nm}$ and water immersion ($n=1.33$) were radially FWHM of $\delta r=0.6\,\lambda/\text{NA}=780\text{ nm}$ and axially $\delta z = 2\,\lambda n/\text{NA}^2=5.0\text{ \textmu m}$. This is consistent with our earlier simulations in Zemax (cf. Fig.~\ref{fig2}), and suggests that our microscope is operating very close to the diffraction limit, and its design specifications.
%
\begin{figure}[t]
    \includegraphics[width=\textwidth]{fig5.jpg}
    \caption{{\bf System capabilities} \textbf{A)} Measurement of the local magnification with a 100\microns calibration target in a thin Fluorescein film below a coverslip. Shown in green is a $700\times700\text{ \textmu m}^2$ square as reference. \textbf{B)} Histogram of the magnifications in A, measured across all squares. \textbf{C)} Two-photon intensity measurements from a uniform bath of Fluorescein. Shown are pixel values (maximum is 32768). \textbf{D)} Horizontal and vertical profiles through the data in C. The green box is indicated by the shaded green area. Range is $\pm 1 \text{ standard deviation}$. \textbf{E)} Histogram of the intensities from within the green box in C. \textbf{F)} Volumetric measurement of a 0.2 \textmu m bead sample in 1\% Agarose. \textbf{G)} Average of N=38 beads. \textbf{H)} Horizontal and vertical profiles through data in G and a Gaussian fit reveal a radial resolution of $760\pm30\text{ nm}$. \textbf{I)} Estimate of the axial resolution shows $5.4\pm0.9\text{ \textmu m}$.}
    \label{fig5}
\end{figure}
%
%
\subsection*{Imaging performance}

@@

For zebrafish imaging, we used the HuC:H2B-GCaMP6f line of transgenic Zebrafish \cite{Cong2017}. Neurons in this organism expresses the fluorescent protein GCaMP6f in their nuclei whose green fluorescence strongly depends on the cellular concentration of Calcium ions. We imaged a 5-day old larva in 2\% low melting point Agarose in E3 medium, see Fig.~\ref{fig7}. We first performed a volumetric scan of the organism, Fig.~\ref{fig7}A shows a rendering in Fiji \cite{Schindelin2019}, and then focused on a plane with a large number of visible neurons, Fig.~\ref{fig7}B. In a small region, we used our piezoelectric collar to collect volumetric data from three planes, see Fig.~\ref{fig7}C. We processed these data with Suite2p \cite{Pachitariu2016}, which identified many neurons with complex calcium transients, Fig.~\ref{fig7}D.\newline

@@

is shown as thin purple line. Deriving this from geometric principles, and aligning the Pockels cell on the table for greatest dynamic range is a relatively straightforward. The students perform these measurement by scanning through voltage settings on the Pockels driver, while monitoring light intensity. This explains to them the particular orientation of the crystal relative to the laser.\newline

@@

Our microscope was set up in a teaching lab space at the Princeton Neuroscience Institute. Over the course of three years, it was disassembled and re-assembled by various generations of students and postdocs. We found that assembling and aligning the microscope, as described, takes around 10 work-days. These two weeks seem to be a good compromise between depth, and researcher commitment. Various additional content across difficulty levels can be added depending on student interest and abilities (cf. Fig.~\ref{fig9}). For example, and only mentioned in passing before, the history and current applications of the two-photon effect and its use in microscopy is a fascinating subject to explore, e.g. \cite{Sheppard2020, Yao2023}.\newline 

@@

\paragraph*{S1 Appendix:}
\label{S1_Appendix}
A document with detailed build instructions and pictures of the microscope.

\paragraph*{\bf Online Repository:} 
\label{repository}
All CAD and optics design files, bill-of-materials, and other useful information: \url{https://github.com/BrainCOGS/Microscope}.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{TWINKLE: An open-source two-photon microscope for teaching and research}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Manuel Schottdorf\textsuperscript{1,*}, % 0000-0002-5468-4255
P. Dylan Rich\textsuperscript{1}, % 0000-0001-9782-7984
E. Mika Diamanti\textsuperscript{1}, % 0000-0003-1199-3362
Albert Lin\textsuperscript{1,4}, % 0000-0002-4541-5889
Sina Tafazoli\textsuperscript{1}, % 0000-0003-1926-0227
Edward H. Nieh\textsuperscript{1,2}, % 0000-0003-2154-6224
Stephan Y. Thiberge\textsuperscript{1,3*} % 0000-0002-6583-6613
\\
\bigskip
\textbf{1} Princeton Neuroscience Institute, Princeton University, Princeton, NJ, USA\\
\textbf{2} School of Medicine, University of Virginia, Charlottesville, VA, USA \\
\textbf{3} Bezos Center for Neural Circuit Dynamics, Princeton University, Princeton, NJ, USA\\
\textbf{4} Center for the Physics of Biological Function, Princeton University, Princeton, NJ, USA\\
\bigskip

@@

\section*{Materials and methods}
In two-photon laser-scanning microscopy\cite{Denk1990}, laser light is focused to a small volume. The product of the interaction of light and sample, fluorescent light, is then measured with a sensitive light detector. Moving the volume of light through the sample and combining the detected fluorescence from different $x$ and $y$ positions produces a picture. The key to two-photon microscopy is the light source: A femtosecond-pulsed infrared laser restricts the volume in which fluorescence occurs to a tiny region in the sample with sufficiently high photon density for the near-simultaneous absorption of two photons by a fluorescent molecule \cite{Denk1990, Svoboda1997, So2000, Helmchen2005}.\newline
Here, we design an open and high-performing microscope that uses these principles. As such we have made all optical and mechanical designs, electronics, bill-of-materials, CAD assembly, testing results, and relevant schematics open to the public. Excluding laser and optical table, the cost is around US-\$ 110k (2024). A narrative of the design and high-level view is presented here. For further technical details, and complete built documentation, we refer the reader to \nameref{S1_Appendix}, and the repository at \url{https://github.com/BrainCOGS/Microscope}.

\subsection*{Design specifications}
To summarize our design goals, we aimed for a mechanical and optical assembly using as many off-the-shelf components as possible, with only few custom aluminium parts that can be machined in any university machine shop. Our design lacks a microscope body, which is  advantageous for two reasons: it aids the use as a teaching instrument, and facilitates adaptation for \invivo imaging because it provides space for the organism and ancillary hardware. We also avoid any custom optics to aid transparency of the function of the optical system. Our system is also made cost-effective, in part, by the availability of femtosecond laser systems based on fiber technology. Operating at a fixed wavelength, these systems come at the fraction of the cost of a more traditional tunable Ti:Sapphire laser \cite{Bueno2019}.\newline
We aimed for (1) $\gtrapprox20\text{ cm}$ of free space around the objective in all directions to aid integration of the microscope into custom behavior boxes, or sample container, and (2) resolution limited imaging of at least $\approx 700\times 700\text{ \textmu m}^2$ field of view which is typical for imaging brain tissue at cellular resolution. (3) Compared to previous open designs\cite{Rosenegger2014, Mayrhofer2015}, our system can make use of the large back aperture of modern long-working-distance objectives to collect more fluorescence and (4) uses GaAsP photomultiplier tubes to maximize signal to noise.\newline
An overview of the microscope is shown in Fig.~\ref{fig1}. Fig.~\ref{fig1}A shows the system as a cartoon. Full details will be provided below. In short, we chose a Spark Alcor 920 nm femtosecond pulsed laser, whose beam, after $2\times$ magnification with a telescope, reaches the scanning mirrors. The scanning mirrors are geometrically close together in the focal plane of the scan lens. This simplifies the design, but leads to minor distortions across the field of view (a few percent at low magnifications, see below). The combined scan and tube lenses magnify the beam further to fill the back aperture of the objective. Fluorescence from the specimen is then collected and focused on the aperture of two GaAsP photomultiplier tubes with suitable color filters. Fig.~\ref{fig1}B shows the CAD design of the entire assembly with important parts of the optical path labelled. Fig.~\ref{fig1}C shows an annotated picture of the microscope head. Notice that compared to the cartoon in Fig.~\ref{fig1}A, the optical path is folded with several mirrors. These mirrors significantly simplify alignment, and make the design more compact.

@@

\subsection*{Assembly on the table}
Various components are used between the laser and the microscope head to condition the beam, and  for intensity control, Fig.~\ref{fig1}B. Out of the laser head, the beam first encounters a waveplate (AHWP10M-980, Thorlabs). This is a birefringent crystal that can rotate the polarization of the laser. Next, the beam travels through a polarizing beamsplitter cube (PBS103, Thorlabs). This splits the beam into two beams whose relative intensity can be adjusted by the orientation of the waveplate. We set the orientation of the waveplate to split the beam into two beams with 50\% power each to feed a second experimental setup. In the CAD drawing, this second beam is sent to a beam block (LB1, Thorlabs). Next, the beam passes through an electro-optic modulator, a Pockels cell (350-80-LA-02, Conoptics). This device allows rapid and electronic control of the laser's power level. It uses a crystal whose refractive index is controlled by an external electric field, combined with a polarizer. This can be thought of as a voltage-controlled wave plate, in which the electric field controls how much light travels through the polarizer. Next, the beam travels through an open mechanical shutter (LS6S2ZM1, Vincent Associates). When closed, the reflection from the shutter is used for calibrating the Pockels cell with a photodiode (PDA36A2, Thorlabs). Following the shutter, the beam travels though a $2\times$ telescope (GBE02-B, Thorlabs). This magnifies the beam diameter by a factor of 2. Past the telescope, the beam is then reflected off several mirrors in a periscope configuration before entering a custom aluminium box housing, mounted to a 95 mm optical rail, that houses the two scanning mirrors (CRS8K/6215H scanning mirror Set, Novanta), shown in blue in  Fig.~\ref{fig1}C. Based on their angles, the two mirrors deflect the beam in x and y position. The beam then travels through scan lens (green) and tube lens (red), the collection optics (yellow) and finally the objective (cyan). The design of these components is covered in the next paragraphs. The steering mirrors in our system are all silver mirrors from Thorlabs. Silver mirrors contribute minimally to dispersion which is inevitable when light travels through various optical components. Limiting the dispersion makes it possible to correct for it with the laser's built-in dispersion compensation.

@@

We assembled the scan and tube lenses as lens groups from off-the-shelf parts, using the same elements as in \cite{Yao2023}, and further optimizing lenses and spacing in Zemax (an optics design software). Details are provided below. After optimization, the focal lengths of scan and tube lens were 99 mm and 378 mm respectively, deviating around 1\% from their target values (Zemax simulation at 920 nm).

@@

    \caption{{\bf Optical design of the excitation pathway.} \textbf{A)} Zemax design with the scanning mirror, tube and scan lens groups. Colors indicate light rays produced by different mirror angles of the scanning mirror. Notice how the rays converge in the image plane. \textbf{B)} CAD design built around the optical design in A, see also Fig.~\ref{fig1}B. \textbf{C)} Cuts through the scan and tube lens assemblies, which are fabricated from off-the-shelf lenses housed in SM2 lens tubes. The specific lenses are stated in the text. \textbf{D)} Simulation of the optical system's two-photon point spread function (PSF-2p) for 0 deg deflection angle, and two example radial sections. \textbf{E)} PSF-2p in the xz plane across the full range of deflection angles. Shown below is Petzval field curvature (blue), and radial (purple) and axial (yellow) full-width-at-half-maximum of the PSF-2p. Error bars are resolution limits of the Huygens PSF estimates in sequential mode.}

@@

Fig.~\ref{fig2}A shows the propagation of light rays for various deflection angles of the scanning mirrors, and how the beams converge on a line under the objective. The Zemax design was then exported into CAD, and the optomechanics assembled around it, see Fig.~\ref{fig2}B. The CAD design and lens groups are highlighted in Fig.~\ref{fig2}B,C (see Fig.~\ref{fig1}C of a photograph of the same part of the instrument). The key property that we optimized when designing the excitation system was an essentially flat, and diffraction limited focal volume across a wide range of deflection angles of the scanning mirrors. Fig.~\ref{fig2}D shows an estimate of the two-photon point spread function (PSF-2p), computed as the square of the Huygens PSF in Zemax. The maximum scan angle of $\pm 10\text{ deg}$ is limited by vignetting of the two inch diameter optics. Analyzing the PSF-2p across optical deflection angles, Fig.~\ref{fig2}E, suggests very small field curvature. In this simulation, the axial resolution was $\delta z = 5.0\text{ \textmu m}$, and the radial resolution $\delta r = 0.75\text{ \textmu m}$ across a wide range of deflection angles.\newline

@@

We incorporated four design principles into our assembly: (1) Fluorescence light must fall approximately perpendicularly on the optical filters for the filters to work correctly. Narrow filters are particularly susceptible to this. Although light emitted from the image plane of an infinity-corrected objective is collimated as it returns from the sample, actual fluorescence weakly diverges due to scattering and diffraction in the material between image plane and objective (e.g. brain tissue \cite{Zipfel2003, Taddeucci1996}). The detector assembly should therefore be placed as close as possible to the objective where fluorescent light will be nearly collimated. This placement reduces sensitivity to spurious light and allows IR light to travel through the collection optics unimpeded, while fluorescent light is reflected into the optical assembly. This arrangement also has the advantage that possible astigmatism, introduced by curvature of the dichroic in the reflected light, only affects the collection optics which does not have to form an image. (2) The collection lens should be large to maximize the collection of as much fluorescent light as possible. This implies that the collection lens should follow the dichroic mirror as closely as possible. (3) Having the collected photons collimated on the aperture of the photomultiplier tubes reduces the sensitivity of the system on surface inhomogeneities of the detector\cite{Tsai2002, Young2015}. In scanning microscopy, an image is produced as a time series of intensities. It is therefore more important to collect a consistently large portion of fluorescent light, rather than to minimize aberration in the collection optics. We therefore optimized relative illumination as function of the field angle in Zemax. (4) All dichroic mirrors and optical band-pass filters have finite performance. It it therefore advantageous to combine multiple band-pass filters to isolate relevant fluorescence ranges. We aimed to record fluorescence from two well-separated spectral ranges, a green ($525\pm25\text{ nm}$) and a red channel ($600\pm25\text{ nm}$). Beyond an IR block in the collection optics, we chose to only use simple long-pass filters, which means that the red channel is operated in the transmission, and the green channel in the reflection mode (cf. Fig.~\ref{fig1}A).\newline

@@

\subsection*{Mechanical design}
As mounting system of our microscope, we chose a standard 8 inch thick optical table. The design was done in CAD, and the results shown in Fig.~\ref{fig1}B. The aim was a mechanical assembly that uses as few custom parts as possible. We discuss this further in the discussion section. The components in the CAD files can be assembled and aligned by an experienced researcher in a few days. When used for teaching, a careful assembly and alignment is viable with about a week of work. Regarding the footprint, the system can comfortably be built on an $4\times4$ ft area of table. The microscope documented here was setup on a $4\times8$ ft table, shared with a second microscope. The beam-splitting hardware is shown in Fig.~\ref{fig1}B as well.\newline
The custom mechanical components (e.g. the light-tight aluminium housing of the collection optics) were built in the university's machine shop from aluminium stock or Thorlabs parts. Some parts have tight tolerances which suggests manufacturing in a professional machine shop is preferred. If useful for training, this could however be done by students. An example are the aluminium parts that hold the large dichroic mirrors and optical filters in the collection optics. After drilling the apertures for light to pass through, not much metal is left on the part which can lead to warping and distortions.\newline
Finally, we want to emphasize the difference between a theoretically optimal design in Zemax and finite tolerances in real-world optomechanical parts. It is key for a good microscope design to be robust against such errors. In our design, the objective is fixed, while the mirrors, the tube and scan lens assemblies, and the scanning mirrors can move along the optical axis for alignment. This allows the mechanical components sufficient degrees of freedom to optically align all key parts of the microscope. In the supplement, \nameref{S1_Appendix}, we go through this alignment procedure in detail.

@@

\subsection*{System performance}
%
Here, we document the performance of the microscope and a few example applications using a Nikon NA 0.8 16$\times$ LWD objective (N16XLWD-PF, Thorlabs). We first determined the size and properties of the field of view with a 100~\textmu m grid (R1L3S3P, Thorlabs), imaged using a Fluorescein in water film through a standard \#1.5 ($\approx170\text{ \textmu m}$ thick) coverslip, see Fig.~\ref{fig5}A. The field of view was $d\approx1.3\text{ mm}$ along the diagonal, $A\approx 1\text{ mm}^2$ in area and the field curvature below the thickness of a thin Fluorescein film $\lesssim 10\text{ \textmu m}$. This is consistent with our earlier Zemax simulations. When measuring the magnification $M$ in \textmu m/px across the field of view, we observed small deviations of $\Delta M \approx 0.1\text{ \textmu m/px}$ in the center when compared to the edges, while the overall range of magnifications across the field of view was tight, see Fig.~\ref{fig5}B. The average magnification is $M=2.77\pm0.04\text{ \textmu m/px}$, which suggests a deviation from a perfect f-Theta system of $\Delta M/M = 0.1/2.8 \approx 4\%$. Next, we imaged a uniform bath of Fluorescein, see Fig.~\ref{fig5}C. This is a combined measure of the quality of excitation and the efficiency of the collection optics across the field of view. Within the central $700\text{ \textmu m} \times 700\text{ \textmu m}$ region, the signal deviated from uniformity within $\approx 13\%$ ($\pm 1 \text{ standard deviation}$) suggesting a relatively flat field, see Fig.~\ref{fig5}D, E. Following these measurements, we imaged a bead sample of 0.2\microns diameter in 1\% Agarose (Dragon Green beads; Bangs Laboratories) to measure axial and radial PSFs. The sample is shown in Fig.~\ref{fig5}F. Averaging across the $N=38$ beads in this volume, we measured a radially symmetric full-width-at-half-maximum of $\delta r\approx760\pm30\text{ nm}$, see Fig.~\ref{fig5}G,H, and axially $\delta z \approx 5.4\pm0.9\text{ \textmu m}$, see Fig.~\ref{fig5}I. The theoretical resolution limits\cite{Tsai2002} for our system, underfilled to $\text{NA}=0.7$ at $\lambda=920\text{ nm}$ and water immersion ($n=1.33$) were radially FWHM of $\delta r=0.6\,\lambda/\text{NA}=780\text{ nm}$ and axially $\delta z = 2\,\lambda n/\text{NA}^2=5.0\text{ \textmu m}$. This is consistent with our earlier simulations in Zemax (cf. Fig.~\ref{fig2}), and suggests that our microscope is operating very close to the diffraction limit, and its design specifications.
%
\begin{figure}[t]
    \includegraphics[width=\textwidth]{fig5.jpg}
    \caption{{\bf System capabilities} \textbf{A)} Measurement of the local magnification with a 100\microns calibration target in a thin Fluorescein film below a coverslip. Shown in green is a $700\times700\text{ \textmu m}^2$ square as reference. \textbf{B)} Histogram of the magnifications in A, measured across all squares. \textbf{C)} Two-photon intensity measurements from a uniform bath of Fluorescein. Shown are pixel values (maximum is 32768). \textbf{D)} Horizontal and vertical profiles through the data in C. The green box is indicated by the shaded green area. Range is $\pm 1 \text{ standard deviation}$. \textbf{E)} Histogram of the intensities from within the green box in C. \textbf{F)} Volumetric measurement of a 0.2 \textmu m bead sample in 1\% Agarose. \textbf{G)} Average of N=38 beads. \textbf{H)} Horizontal and vertical profiles through data in G and a Gaussian fit reveal a radial resolution of $760\pm30\text{ nm}$. \textbf{I)} Estimate of the axial resolution shows $5.4\pm0.9\text{ \textmu m}$.}
    \label{fig5}
\end{figure}
%
%
\subsection*{Imaging performance}

@@

For zebrafish imaging, we used the HuC:H2B-GCaMP6f line of transgenic Zebrafish \cite{Cong2017}. Neurons in this organism express the fluorescent protein GCaMP6f in their nuclei whose green fluorescence strongly depends on the cellular concentration of Calcium ions. We imaged a 5-day old larva in 2\% low melting point Agarose in E3 medium, see Fig.~\ref{fig7}. We first performed a volumetric scan of the organism, Fig.~\ref{fig7}A shows a rendering in Fiji \cite{Schindelin2019}, and then focused on a plane with a large number of visible neurons, Fig.~\ref{fig7}B. In a small region, we used our piezoelectric collar to collect volumetric data from three planes, see Fig.~\ref{fig7}C. We processed these data with Suite2p \cite{Pachitariu2016}, which identified many neurons with complex calcium transients, Fig.~\ref{fig7}D.\newline

@@

is shown as thin purple line. Deriving this from geometric principles, and aligning the Pockels cell on the table for greatest dynamic range is relatively straightforward and the students perform these measurement by scanning through voltage settings on the Pockels driver, while monitoring light intensity. This explains to them the optimal orientation of the crystal relative to the laser.\newline

@@

Our microscope was set up in a teaching lab space at the Princeton Neuroscience Institute. It was disassembled and re-assembled by various generations of students and postdocs. We found that assembling and aligning the microscope, as described, takes around 10 work-days. These two weeks seem to be a good compromise between depth, and researcher commitment. Various additional content across difficulty levels can be added depending on student interest and abilities (cf. Fig.~\ref{fig9}). For example, and only mentioned in passing before, the history and current applications of the two-photon effect and its use in microscopy is a fascinating subject to explore, e.g. \cite{Sheppard2020, Yao2023}.\newline 

@@

\paragraph*{S1 Appendix}
\label{S1_Appendix}
A document with detailed build instructions and pictures of the microscope.

\paragraph*{\bf Online Repository}
\label{repository}
All CAD and optics design files, bill-of-materials, and other useful information: \url{https://github.com/BrainCOGS/Microscope}.
